---
title: "Dominant palmetto species (*Serenoa repens* and *Sabal etonia*) in Archbold Biological Station, FL"
author: "Teague Tran"
date: "2/7/2021"
output: 
  html_document:
    theme: cerulean
    code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)
library(lubridate)
library(dplyr)
library(patchwork)
library(GGally)
library(broom)
library(jtools)
library(kableExtra)
library(scales)

# read in data
palmetto <- read_csv(here("data", "palmetto.csv"))
```

## Data Visualization/Exploration

```{r, warning = FALSE, message = FALSE}
# select species, height, length, width, and green leaf count

selectedattributes_pal <- palmetto %>% 
  select(species, height, length, width, green_lvs, year) %>% 
  mutate(species_names = 
           case_when(
             species == "1" ~ "Serenoa repens",
             species == "2" ~ "Sabal etonia")
         ) 
# find the mean height by each species
mean_heights <- selectedattributes_pal %>%
  group_by(species_names) %>% 
  summarize(
    mean_height = mean(height, na.rm = TRUE),
    sd_height = sd(height, na.rm = TRUE)
  )

#group by species and year to calculate mean leaf count per recorded year
mean_leaves <- selectedattributes_pal %>%
  group_by(species_names, year) %>% 
  summarize(
    mean_leaves_count = mean(green_lvs, na.rm = TRUE)
  )
#group by species and calculate mean length and width to put in graph
mean_lw <- selectedattributes_pal %>% 
  group_by(species_names) %>% 
  summarize(
    mean_length = mean(length, na.rm = TRUE),
    sd_length = sd(length, na.rm = TRUE),
    mean_width = mean(width, na.rm = TRUE), 
    sd_width = sd(width, na.rm = TRUE)) 

```
```{r, warning = FALSE, message = FALSE}
#Explore the differences between the two species 

#visualize mean height difference between the two species
palmetto_heights <- ggplot(data = mean_heights, aes(x = species_names, y = mean_height)) +
  geom_col(fill = c("coral","navy")) + 
  theme_minimal() +
  labs(x = " ", y = "mean height (cm)", title = "Mean height of two palmetto species in Archbold Biological Station, FL", subtitle = "(1981-2017)")

#visualize mean leaf count difference between the two species
palmetto_leaves <- ggplot(data = mean_leaves, aes(x = year, y = mean_leaves_count, color = species_names)) +
  geom_point(size = 2) +
  theme_minimal() + 
  scale_color_manual(values = c("coral", "navy")) +
  labs(x = "Year", y = "Mean leaf count (# leaves)", title = "Mean leaf count of two palmetto species in Archbold Biological Station, FL", subtitle = "(1981-2017)", color = "Species Names")

#visualize the mean of width between the two species

palmetto_length <- ggplot(data = mean_lw, aes(x = species_names, y = mean_length)) + 
  geom_col() +
  geom_col(fill = c("coral","navy")) + 
  theme_minimal() +
  labs(x = " ", y = "mean length (cm)", title = "Mean canopy length of two palmetto species in Archbold Biological Station, FL", subtitle = "(1981-2017)")

#visualize the mean of length between the two species
palmetto_width <- ggplot(data = mean_lw, aes(x = species_names, y = mean_width)) +
  geom_col() + 
  geom_col(fill = c("coral","navy")) + 
  theme_minimal() +
  labs(x = " ", y = "mean width (cm)", title = "Mean canopy width of two palmetto species in Archbold Biological Station, FL", subtitle = "(1981-2017)")
#patch the width and length graphs together
palmetto_lwh_combined <- (palmetto_width / palmetto_length)

```

```{r, warning = FALSE, message = FALSE}
#display leaves grpah with figure caption under
palmetto_leaves 
```

Figure 1. *Serenoa repens* consistently has a higher count of leaves than *Sabal etonia* every year data was recorded. 

```{r, warning = FALSE, message = FALSE}
#display heights grpah with figure caption under
palmetto_heights
```

Figure 2. Mean height of *Sabal etonia* (`r round(mean_heights$mean_height[1], 2)` $\pm$ `r round(mean_heights$sd_height[1], 2)`) is lower than *Serenoa repens* (`r round(mean_heights$mean_height[2], 2)` $\pm$ `r round(mean_heights$sd_height[2], 2)`). 

```{r,warning = FALSE, message = FALSE}
#display patched graph with figure caption under
palmetto_lwh_combined
```

Figure 3. Both mean canopy width and mean canopy length of *Sabal etonia* (`r round(mean_lw$mean_width[1], 2)` $\pm$ `r round(mean_lw$sd_width[1], 2)` and `r round(mean_lw$mean_length[1], 2)` $\pm$ `r round(mean_lw$sd_length[1], 2)` respectively) are higher than the mean canopy width and mean canopy length of *Serenoa repens* (`r round(mean_lw$mean_width[2], 2)` $\pm$ `r round(mean_lw$sd_width[2], 2)` and `r round(mean_lw$mean_length[2], 2)` $\pm$ `r round(mean_lw$sd_length[2], 2)` respectively).

## Binary Logistic Regression


```{r, warning = FALSE, message = FALSE}
# attributes already selected in selectedattributes_pal from part 1
binlog_pal <- selectedattributes_pal %>% 
  mutate(species_names = fct_drop(species_names)) 
# check levels of factors; Sabal etonia = 0, Serenoa repens = 1
# Used levels to check; used "#" to not show levels(binlog_pal$species_names)

#perform binary logistic regression
blr_pal <- glm(species_names ~ height + length + width + green_lvs,
               data = binlog_pal,
               family = "binomial")

#let's take a look at the blr! 
# blr_pal

# summary(blr_pal)

### commented out to not include in data

#get tidy version 

blr_tidy <- broom::tidy(blr_pal)

#remove intercept row from blr_tidy
blr_tidier <- blr_tidy[-c(1),] %>% 
  mutate(term = case_when(
            term == "height" ~ "Height",
            term == "width" ~ "Width", 
            term == "length" ~ "Length",
             term == "green_lvs" ~ "Green leaf count"))
  
blr_tidier <- blr_tidier %>% 
  mutate(p.value = case_when(
            p.value < .05 ~ "< .05"))

```
```{r, warning = FALSE, message = FALSE}
#create a 
blr_tidier %>% 
  kable(col.names = c("Variable", "Coefficient", "Standard Error", "Z-statistic", "P-value" ), digits = 5) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```


```{r, warning = FALSE, message = FALSE}
#evaluate success of this model (>=50% is cutoff)
blr_fitted <- blr_pal %>% 
  augment(type.predict = "response")
#select and compare predicted species vs actual species
blr_fitted <- blr_fitted %>% 
  mutate(prediction = 
          case_when(.fitted >= .5 ~ "Serenoa repens",
                    .fitted < .5 ~ "Sabal etonia"))
blr_fitted <- blr_fitted %>% 
  mutate(correct = 
           case_when(prediction != species_names ~ "incorrect",
                     prediction == species_names ~ "correct"))
#set data up to be ready to be put in a table
blr_table <- blr_fitted %>% 
  select(species_names, correct) %>% 
  group_by(species_names, correct) %>% 
  count() %>% 
  pivot_wider(
    names_from = correct,
    values_from = n)
blr_table <- blr_table %>% 
  mutate(percentage = percent(correct/(incorrect + correct)))

#create table using Kable

blr_table %>% 
  kable(col.names = c("Species", "Number of correct", "Number incorrect", "Percentage Correct")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
  
```


## Citation: 

Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative. [https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5](https://doi.org/10.6073/pasta/f2f96ec76fbbd4b9db431c79a770c4d5)






































